The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale

### 2022-08-11 manuscript/main.tex
update factor loading thresholds; updates to manuscript (edits + reformatting floats)

\documentclass[letterpaper,man,natbib,floatsintext,longtable]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}             % fancy latex tables
\usepackage[export]{adjustbox}    % center wide tables on page
\usepackage{setspace}             % adjust caption linespacing 
\usepackage{multibib}             % references in supplement

@@

@@

\textbf{Background}

\textbf{Methods}

\textbf{Results}

\textbf{Conclusions}

}

@@

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}. Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Thus, reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages. The MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment. Although six of these subscales overlap with other standard measures of abuse (e.g. physical abuse, neglect), the remaining four are less commonly represented in other scales (e.g. peer victimization, witnessing domestic violence). The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposures. The MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development. The MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced). These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}. For these reasons the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores. For example, some have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Inspired by a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made new subscale scores indexing threat and deprivation experiences using the items in the MACE that respectively measure abuse and neglect \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}. The psychometric properties of these new subscale scores have not been thoroughly investigated.

A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of adverse experiences. That is, children who have experienced any single form of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}. As a consequence, isolating the unique contribution of one type of maltreatment to some outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment. Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they draw spurious conclusions.

Bifactor models, a type of hierarchical factor model, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. They assume that the covariation in responses to a set of items can be accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors. Applied to the MACE, these indices can help determine whether subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead chiefly reflect overall maltreatment.

summary of paper paragraph tk. data availability statement to conclude

@@

\subsection{Participants}

@@

\subsection{Measures}

@@

\subsection{Exclusion criteria}

@@

\subsection{Differential item functioning}

@@

\subsection{Response dependence}

@@

\subsection{Confirmatory item factor models}

@@

\subsection{Goodness of fit \& model comparison}

@@

\subsection{Model-based reliability indices}

To evaluate the reliability of the MACE total and subscale scores, we calculated several model-based reliability indices using the factor loadings from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. both general and specific factors). Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}). Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor. The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score. When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor \citep{rodriguez2016applying}. Values of $\omega_s$ approaching 1 indicate the specific factor, not the general factor, is the primary source of reliable variance in a subscale score. \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, values of $\omega_s$ < 0.50 preclude the interpretation of subscale scores as primarily reflecting a specific factor \citep{gignac2013bifactor}.

@@

\subsection{Exploratory factor analysis}

In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data from each sample separately. The goal of the EFA was to examine what factor structure would emerge with no \emph{a priori} restrictions placed on the data. We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models. We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (maladaptive parenting, peer victimization, reverse-scoring).

@@

The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the data. All RMSEA value were <0.05 and most CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data where CFI/TLI > 0.90. In general, model fits were better for the replication sample data than for the original sample. Furthermore, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical values indicating that all models were able to reproduce the observed distribution of total scores for each sample.

\begin{table}[t!]
\small
\centering
\begin{adjustbox}{center}
\begin{tabular}{ccccccccccr}
\toprule
Sample & Model & $\chi^2 (df)$ &  RMSEA & CFI & TLI & $\chi^2_{NC}$ (PPP) & $Q3_{\max}$ &  $Q3_{\text{crit}}$ & LOO-CV & \multicolumn{1}{c}{$\Delta$LOO (se)} \\
\midrule
Original & 1 &  1858 (702) &  0.040 &  0.927 &  0.923 &  0.051 (0.230) &  0.589 &  0.239 &  30967.2 &  1748.0 (76.7) \\
& 2 &  1373 (663) &  0.032 &  0.955 &  0.950 &  0.046 (0.356) &  0.573 &  0.300 &  29211.3 & \multicolumn{1}{c}{-} \\
& 3 &  1571 (692) &  0.035 &  0.944 &  0.940 &  0.051 (0.214) &  0.591 &  0.255 &  30531.1 &  1312.0 (68.0) \\
& 4 &  1253 (687) &  0.028 &  0.964 &  0.961 &  0.046 (0.346) &  0.591 &  0.279 &  29219.2 & 7.9 (56.3) \\
\midrule
Replication & 1 &  1086 (702) &  0.031 &  0.954 &  0.952 &  0.065 (0.629) &  0.513 &  0.250 &  19771.6 &  1089.3 (58.4) \\
& 2 &   956 (663) &  0.028 &  0.965 &  0.961 &  0.058 (0.766) &  0.456 &  0.310 &  18706.2 &    24.0 (52.0) \\
& 3 &   881 (692) &  0.022 &  0.977 &  0.976 &  0.062 (0.684) &  0.468 &  0.316 &  19177.7 &   495.5 (45.0) \\
& 4 &   842 (687) &  0.020 &  0.982 &  0.980 &  0.058 (0.756) &  0.437 &  0.318 &  18682.2 & \multicolumn{1}{c}{-} \\
\bottomrule
\end{tabular}
\end{adjustbox}
\captionsetup{width=1.1\textwidth}
\caption{\normalfont Fit statistics for the confirmatory factor models by sample. LOO-CV values are presented in deviance scale (i.e. smaller values indicate better fit). Notes: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
\label{table:cfa_diagnostics}
\end{table}

@@

\subsection{Confirmatory item response models}

@@

\subsubsection{Model 1: One-factor (unidimensional) model of the MACE}

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
    \captionsetup{width=1.1\textwidth}
    \caption{Standardized factor loadings from the four models estimated from the original sample. Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.  ** Reverse-scored items.}
    \label{fig:loadings_original}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
    \captionsetup{width=1.1\textwidth}
    \caption{Standardized factor loadings from the four models estimated from the online sample. Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. ** Reverse-scored items.}
    \label{fig:loadings_online}
\end{figure}

\subsubsection{Model 2: Bifactor model with 10 specific factors}

@@

\subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}

@@

\subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}

@@

\subsection{Exploratory factor analysis}

\begin{figure}[t!]
    \centering
    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row). Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. Only factor loadings $\geq$ 0.30 are shown. ** Reverse-scored items.}
    \label{fig:efa_cf}
\end{figure}

@@

@@

A second limitation is our sample demographics. Though the replication sample increased the overall diversity of our sample with respect to gender and age, other types of diversity are notably lacking. Our combined sample was neither racially nor ethnically diverse. This is important as previous studies have identified DIF for childhood adversity questionnaires by race \citep{rodriguez2019identification}, which raises the question whether similar issues would be found for the MACE in more diverse samples. Moreover, we are not currently able to explain why we found DIF for multiple  sample membership in the current study. Additional research is needed to study the functioning of the MACE in more diverse samples in order to validate its use in different populations.

To conclude, the results of the current study show the factor structure of the MACE is essentially unidimensional. Bifactor models revealed that, despite the presence of smaller secondary factors, the majority of reliable variance in the MACE endorsement responses are attributable to a general maltreatment factor. Accordingly, we found that the MACE total score is a reliable and valid of overall maltreatment. In contrast, we found that MACE subscale scores are not reliable measures of particular types of childhood maltreatment; instead they mainly reflect general maltreatment. Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should ensure that any subscale scores used as part of their analyses are reliable, lest they draw spurious inferences.

@@

%For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model. The vast majority of items on the MACE measure maltreatment by parents or other adults in the home. Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g. maladaptive parenting, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor. Moreover, there is theoretical support for a model in which a common source like maladaptive parenting, itself caused by a diversity of parent level (e.g. emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}. This is empirically supported insofar that multiple forms of maltreatment share overlapping risk factors \citep{stith2009risk, mulder2018risk, assink2019risk}. Furthermore, the MACE is a retrospective self-report measure. Prospective and retrospective measures do not have perfect agreement \citep{patten2015retrospective, reuben2016lest}, and thus must reflect other factors related to the participants including individual differences in current state, memory, appraisal, personality, and biases in discloure \citep{susser2012still, reuben2016lest, colman2016consistency}.

%This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor. At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter. However, there is ample evidence that children who have been maltreated are at greater risk for subsequent peer victimization \citep{bolger2001developmental, benedini2016cycle}. The reasons for this association are manifold. For example, maltreatment may influence a child's sociocognitive development (e.g. attachment style, emotion regulation) in ways that may predispose them to peer victimization \citep{goemans2021child, McCrory2022-cj}. Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place. Insofar that maladaptive parenting is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor. This is also highlights the need for future study of the peer victimzation factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.

>>>>

\documentclass[letterpaper,man,natbib,noextraspace,floatsintext,longtable]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}             % fancy latex tables
\usepackage[export]{adjustbox}    % center wide tables on page
\usepackage{setspace}             % adjust caption linespacing 
\usepackage{multibib}             % references in supplement
\usepackage{hyperref}             % manage hyperlink colors

@@

% change hyperlink colors
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}

@@

\textbf{Background}: Typically measures of childhood maltreatment collect information about several types of abuse and maltreatment. The subscales of the retrospective Maltreatment and Abuse Chronology of Exposure (MACE) scale measure 10 such types. Since different types of maltreatment often co-occur, it is not clear whether subscale measures can be used to reflect constructs unique from from overall maltreatment.

\hfill

\textbf{Methods}: Here, we investigated the factor structure of the MACE scale in two independent samples (N=1051 \& N=582). We used both confirmatory and exploratory item response modeling.

\hfill

\textbf{Results}: We found the MACE total score to be a reliable and valid measure of general maltreatment. Confirmatory item response models provided strong evidence for unidimensionality of the MACE, with the exception of a potential peer victimization factor. That is, most items in the MACE load on general factor and apart from peer victimization very little variance was explained by other subscales. Exploratory factor modeling further supported these findings.

\hfill

\textbf{Conclusions}: The MACE is essentially a unidimensional scale. Apart from a separate peer victimization factor, researchers should refrain from using subscales data separately since they mostly reflect overall maltreatment.  
\hfill \break
}

\keywords{child maltreatment, reliability, item response theory, bifactor model}

\authornote{This work was funded in part by an NSF Graduate Research Fellowship. The authors have no conflicts of interest to declare.}

@@

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life including the development of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}. Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}. Thus, reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages. The MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment. Six of these subscales overlap with other standard measures of abuse (e.g. physical abuse, neglect), whereas the remaining four are less commonly represented in other scales (e.g. peer victimization, witnessing domestic violence). The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposure levels. The MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development. The MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced). These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}. For these reasons, the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. In response, some researchers have begun to use the MACE to make new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores. For example, some have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Inspired by a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made new subscale scores indexing threat and deprivation experiences using the items in the MACE that respectively measure abuse and neglect \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}. Yet the psychometric properties of these new subscale scores have not been thoroughly investigated.

A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of these adverse experiences. That is, children who have experienced any single form of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}. As a consequence, isolating the unique contribution of one type of maltreatment to some outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment. Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they risk drawing spurious conclusions.

Bifactor models, a type of hierarchical factor model, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. They assume that the covariation in responses to a set of items can be accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors. Applied to the MACE, these indices can help determine if subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead primarily reflect overall maltreatment.

In this study, we investigated the factor structure of the MACE in two independent samples of participants. First, we tested the fits of a series of confirmatory item response models to the maltreatment data from each sample. These included several bifactor models, which allowed us to quantify the proportion of variance in the data attributable to general maltreatment versus specific types of abuse. We found that the MACE is essentially unidimensional; that is, the majority of reliable variance in participants' responses was attributable to a general maltreatment factor. We then calculated the reliability of the MACE total score and a host of subscale scores. We found the total score was a reliable and valid measure of general maltreatment, whereas most subscale scores contained little-to-no variance unique from general maltreatment. The only possible exception to this trend was for a novel peer victimization subscale. Finally, we performed exploratory factor analysis to examine the factor structure of the MACE without \textit{a priori} restrictions. These analyses supported the results of the confirmatory factor analyses.

@@

All data, code, and analysis materials are publicly available at \url{https://github.com/szorowi1/mace-irt}.

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Participants}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Measures}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}}  \subsection{Exclusion criteria}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Differential item functioning}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Response dependence}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Confirmatory item factor models}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Goodness of fit \& model comparison}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Model-based reliability indices}

To evaluate the reliability of the MACE total and subscale scores, we calculated several model-based reliability indices using the factor loadings from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald2013test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. both general and specific factors). Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}). Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor. The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score. When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor \citep{rodriguez2016applying}. Values of $\omega_s$ approaching 1 indicate the specific factor, not the general factor, is the primary source of reliable variance in a subscale score. \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, values of $\omega_s$ < 0.50 preclude the interpretation of subscale scores as primarily reflecting a specific factor \citep{gignac2013bifactor}.

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Exploratory factor analysis}

In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data from each sample separately. The goal of the EFA was to examine what factor structure would emerge with no \emph{a priori} restrictions placed on the data. We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models. We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (parental maltreatment, peer victimization, reverse-scoring).

@@

The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the data. All RMSEA value were <0.05 and most CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data where CFI \& TLI > 0.90. In general, model fits were better for the replication sample data than for the original sample. Furthermore, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical values indicating that all models were able to reproduce the observed distribution of total scores for each sample.

\begin{table}[t!]
\small
\centering
\begin{adjustbox}{center}
\begin{tabular}{ccccccccccr}
\toprule
Sample & Model & $\chi^2 (df)$ &  RMSEA & CFI & TLI & $\chi^2_{NC}$ (PPP) & $Q3_{\max}$ &  $Q3_{\text{crit}}$ & LOO-CV & \multicolumn{1}{c}{$\Delta$LOO (se)} \\
\midrule
Original & 1 &  1858 (702) &  0.040 &  0.927 &  0.923 &  0.051 (0.230) &  0.589 &  0.239 &  30967.2 &  1748.0 (76.7) \\
& 2 &  1373 (663) &  0.032 &  0.955 &  0.950 &  0.046 (0.356) &  0.573 &  0.300 &  29211.3 & \multicolumn{1}{c}{-} \\
& 3 &  1571 (692) &  0.035 &  0.944 &  0.940 &  0.051 (0.214) &  0.591 &  0.255 &  30531.1 &  1312.0 (68.0) \\
& 4 &  1253 (687) &  0.028 &  0.964 &  0.961 &  0.046 (0.346) &  0.591 &  0.279 &  29219.2 & 7.9 (56.3) \\
\midrule
Replication & 1 &  1086 (702) &  0.031 &  0.954 &  0.952 &  0.065 (0.629) &  0.513 &  0.250 &  19771.6 &  1089.3 (58.4) \\
& 2 &   956 (663) &  0.028 &  0.965 &  0.961 &  0.058 (0.766) &  0.456 &  0.310 &  18706.2 &    24.0 (52.0) \\
& 3 &   881 (692) &  0.022 &  0.977 &  0.976 &  0.062 (0.684) &  0.468 &  0.316 &  19177.7 &   495.5 (45.0) \\
& 4 &   842 (687) &  0.020 &  0.982 &  0.980 &  0.058 (0.756) &  0.437 &  0.318 &  18682.2 & \multicolumn{1}{c}{-} \\
\bottomrule
\end{tabular}
\end{adjustbox}
\captionsetup{width=1\textwidth}
\caption{\normalfont Fit statistics for the confirmatory factor models by sample. LOO-CV values are presented in deviance scale (i.e. smaller values indicate better fit). Notes: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
\label{table:cfa_diagnostics}
\end{table}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Confirmatory item response models}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 1: One-factor (unidimensional) model}

\begin{figure}[tp]
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
    \captionsetup{width=1.1\textwidth}
    \caption{Standardized factor loadings from the four models fit to the original sample. Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scoring); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. Factor loadings $<$ 0.3 are displayed in gray; factor loadings $\geq$ 0.6 are bolded.  ** Reverse-scored items.}
    \label{fig:loadings_original}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
    \captionsetup{width=1.1\textwidth}
    \caption{Standardized factor loadings from the four models fit to the replication sample. Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scoring); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. Factor loadings $<$ 0.3 are displayed in gray; factor loadings $\geq$ 0.6 are bolded.  ** Reverse-scored items.}
    \label{fig:loadings_online}
\end{figure}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 2: Bifactor model with 10 specific factors}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}

@@

{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Exploratory factor analysis}

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row). \break \small Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. Only factor loadings $\geq$ 0.30 are shown. ** Reverse-scored items.}
    \label{fig:efa_cf}
\end{figure}

@@

A second limitation is our sample demographics. Though the replication sample increased the overall diversity of our sample with respect to gender and age, other types of diversity are notably lacking. Our combined sample was neither racially nor ethnically diverse. This is important as previous studies have identified DIF for childhood adversity questionnaires by race \citep{rodriguez2019identification}, which raises the question whether similar issues would be found for the MACE in more diverse samples. Moreover, we are not currently able to explain why we found DIF in multiple items by sample membership in the current study. Additional research is needed to study the functioning of the MACE in more diverse samples in order to validate its use in different populations.

To conclude, the results of the current study show the factor structure of the MACE is essentially unidimensional. Bifactor models revealed that, despite the presence of smaller secondary factors, the majority of reliable variance in the MACE endorsement responses are attributable to a general maltreatment factor. Accordingly, we found that the MACE total score is a reliable and valid of overall maltreatment. In contrast, we found that MACE subscale scores are not reliable measures of particular types of childhood maltreatment; instead they mainly reflect general maltreatment. Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should ensure that any subscale scores used as part of their analyses are reliable, lest they risk drawing spurious inferences.

@@

%For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model. The vast majority of items on the MACE measure maltreatment by parents or other adults in the home. Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g. parental maltreatment, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor. Moreover, there is theoretical support for a model in which a common source like parental maltreatment, itself caused by a diversity of parent level (e.g. emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}. This is empirically supported insofar that multiple forms of maltreatment share overlapping risk factors \citep{stith2009risk, mulder2018risk, assink2019risk}. Furthermore, the MACE is a retrospective self-report measure. Prospective and retrospective measures do not have perfect agreement \citep{patten2015retrospective, reuben2016lest}, and thus must reflect other factors related to the participants including individual differences in current state, memory, appraisal, personality, and biases in discloure \citep{susser2012still, reuben2016lest, colman2016consistency}.

%This in turn raises another question: whether it is reasonable for indicators of parental maltreatment and peer victimization on the MACE to share a common factor. At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter. However, there is ample evidence that children who have been maltreated are at greater risk for subsequent peer victimization \citep{bolger2001developmental, benedini2016cycle}. The reasons for this association are manifold. For example, maltreatment may influence a child's sociocognitive development (e.g. attachment style, emotion regulation) in ways that may predispose them to peer victimization \citep{goemans2021child, McCrory2022-cj}. Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place. Insofar that parental maltreatment is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor. This is also highlights the need for future study of the peer victimization factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
